Spectroscopic confirmation of very-wide stellar binaries and large-separation comoving pairs from \textsl{Gaia DR1}

### 2017-10-09 paper/comoving-followup.tex
overhaul results and figures

\begin{figure}[!ht]
  \begin{center}
    \includegraphics[width=0.85\linewidth]{rv-comparison.pdf}
  \end{center}
  \caption{%
    Comparison of radial velocities measured from our observations with radial
    velocities obtained from the literature search described in
    \sectionname~\ref{sec:rv-validation}.
    Left panel shows our values on the $x$-axis, ${\rm RV}$, against the
    previous literature values on the $y$-axis, ${\rm RV}_{\rm lit}$, for 155
    observed stars (mainly calibration targets) with previously measured radial
    velocities.
    Right panel shows a histogram of the differences, $\Delta{\rm RV} = {\rm RV}
    - {\rm RV}_{\rm lit}$.
    The estimated standard-deviation of the difference distribution implies a
    global systematic uncertainty of $\approx 20~\kms$.
    \label{fig:compare-previous}}
\end{figure}

\section{Methods: Comoving pair probabilities}\label{sec:methods}

@@

Of the 323 observed star pairs, 12 have at least one member with an unreliable
(flagged) radial velocity measurement from the spectroscopic reduction pipeline.
For the remaining 311 pairs, we construct a statistical model to determine the
probability that a given candidate comoving pair is comoving in 3D velocity
given the radial velocity and uncertainty measured in this work combined with
\tgas\ astrometric data.
For this inference, the data are the proper motions in the ICRS reference frame,
$\bs{\mu} = (\mu_{\alpha^*}, \mu_\delta)$,\footnote{$\mu_{\alpha^*} = \mu_\alpha
\, \cos\delta$} and parallax, $\varpi$, (from the
\tgas\ catalog) and barycentric radial velocity, $v_r$, for each star in a
candidate comoving pair; we assume that the sky position for each star $(\alpha,
\delta)$ is known perfectly.
For brevity in the expressions below, we define $\bs{\eta} = (\mu_{\alpha^*},
\mu_\delta, v_r)$, where, e.g., $\bs{\eta}_1$ refers to the proper motion and
radial velocity data for star 1 in the pair.
To simplify some expressions, we additionally ignore any correlations between
the observed parallax and the proper motion components (but keep the covariance
between the two proper motion components).

@@

At the smallest tangential separations of our candidate pairs, ($s_{\rm tan}
\approx 10^{-3}~\pc$), if the two stars are bound, the expected velocity
difference is a few $\kms$ (depending on mass).
We therefore set $\sigma_{\rm co} = 2~\kms$

We generate samples in $f$ using \package{emcee} (\citealt{Foreman-Mackey:2013})
with 28 walkers, keeping track of the value of the marginal likelihood under
each mixture model component for each sample.
We run the sampler for 256 burn-in steps from a small ball of initial positions
centered on $f=0.5$, then restart the walkers from the final position and run
for an additional 1024 steps.
With the samples in $f$ and the mixture model component likelihoods, we compute
the per-pair probabilities of belonging to the comoving component (see the final
section of \citealt{Foreman-Mackey:2014}).

% TODO: summary of assumptions:
% - parallax-proper-motion covariances small
% - Same velocity dispersion for all ages. Could have an age-dependent
%   dispersion
% - Isotropic field

\section{Old section on FMLs}
We compute two fully marginalized likelihoods (FMLs) for the data: (1) assuming
that the two stars are comoving with identical 3D velocity vectors (within some
allowed small dispersion), and (2) assuming that the two stars have independent
3D velocity vectors drawn from a prior velocity distribution.
We set the small allowed dispersion ($s$, in previous work) to $1~\kms$, and use
a mixture of three Gaussian velocity components for the prior, meant to
represent the thin disk, thick disk, and halo velocity distributions relative to
the sun (see \citealt{Oh:2017}).

To compute the new likelihood ratio, we modify the expressions from
\citet{Oh:2017} to include the radial velocity measurements.
In the same notation, this requires the following changes:
\begin{itemize}
  \item The projection matrix, $\mat{M}$, now has to include the radial unit
    vector
    \begin{equation}
      \mat{M} =
        \left(
          \begin{array}{c c c}
            -\sin\alpha & \cos\alpha & 0 \\
            -\sin\delta \, \cos\alpha & -\sin\delta \, \sin\alpha & \cos\delta\\
            \cos\delta \, \cos\alpha & \cos\delta \, \sin\alpha & \sin\delta
          \end{array}
        \right)
    \end{equation}
    where $(\alpha, \delta)$ are the ICRS coordinates (right ascension and
    declination) of a given star.
  \item The data-space velocity vector, $\vec{y}$, for one star now becomes
    \begin{equation}
      \vec{y} =
        \transpose{\left(
          \begin{array}{c c c}
            r_i\,\mu_{\alpha}^* &
            r_i\,\mu_{\delta} &
            v_{r}
          \end{array}
        \right)}
    \end{equation}
    where $\mu_\alpha^*$ is the proper motion in right ascension (including the
    $\cos{\delta}$ term), $\mu_\delta$ is the proper motion in declination, and
    $v_r$ is the radial velocity.
  \item The covariance matrix becomes the block-diagonal matrix
    \begin{equation}
      \mat{\Sigma} = \left(
        \begin{array}{c c}
          r^2 \, \mat{C} & 0 \\
          0 & \sigma_{v_r}^2
        \end{array}
      \right)
    \end{equation}
    where $r$ is the distance, $\mat{C}$ is the proper motion covariance matrix
    (from the \tgas\ catalog), and $\sigma_{v_r}$ is the radial velocity
    uncertainty.
\end{itemize}
To compute the FML for hypothesis 1 for a pair of stars $(i,j)$, the data-space
vector must become the 6-vector
\begin{equation}
  \vec{y} = \transpose{\left(
    \begin{array}{c c}
      \vec{y}_i &
      \vec{y}_j
    \end{array}
    \right)}
\end{equation}
and the covariance matrix becomes the block-diagonal matrix
\begin{equation}
  \mat{\Sigma} = \left(
    \begin{array}{c c}
      \mat{\Sigma}_i & 0 \\
      0 & \mat{\Sigma}_j
    \end{array}
  \right) \quad .
\end{equation}

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=0.5\linewidth]{llr-vs-dv.pdf}
  \end{center}
  \caption{%
    Point estimate of the 3D velocity difference between two stars in a comoving
    pair, $|\bs{v}_1 - \bs{v}_2|$, plotted against the change in
    log-likelihood-ratio computed when including radial velocity data.
    Error bars are shown for a few randomly-chosen, representative pairs and
    represent the 15th and 85th percentiles of velocity differences computed
    from samples from the error distributions of parallax, proper motion
    components, and radial velocity.
    The likelihood ratio increases (i.e. the belief that the stars are comoving
    becomes stronger) when the 3D velocity difference is small.
    \label{fig:llr}}
\end{figure}

If the new likelihood ratio is larger than that computed from astrometry alone,
the added radial velocity information supports the belief that the two stars are
truly comoving in 3-space.
Conversely, if the new likelihood ratio is smaller, the radial velocity data
indicate that the pair is likely a false-positive and are not truly comoving.
In practice, we compute the logarithm of the likelihood ratios,
$\mathcal{R} = \ln \mathcal{L}_1/\mathcal{L}_2$; we refer to the
proper-motion-only log-FML-ratio for a given pair of stars as
\llrold\ (as computed in \citealt{Oh:2017}) and the log-FML-ratio including the
radial velocity data as \llrnew.
We consider a pair as being genuinely comoving if $\llrnew > \llrold$:
\ncomoving\ of the \npairsobs\ comoving pairs satisfy this constraint.
\figurename~\ref{fig:llr} shows the magnitude of the 3D velocity difference
plotted against the change in log-likelihood-ratio, $\llrnew - \llrold$, for all
observed pairs.
In detail, for visualization, we sample from the error distributions over proper
motion, parallax, and radial velocity for each star individually and compute the
magnitude of the 3D velocity differences, $|\bs{v}_1-\bs{v}_2|$, for $2^{16}$
samples.
We plot the median of the resulting distribution of velocity-differences as
dark markers (circles).
The error bars show the 15th and 85th percentiles of the distributions for a few
randomly chosen pairs.
Note that pairs with large 3D velocity differences have large (negative) changes
to their log-likelihood-ratio values.
Some pairs with large velocity differences also have small (near zero) changes
to the log-likelihood ratio: these are pairs where each member has a velocity
that is an outlier with respect to the prior velocity distribution.

The inferred 3D velocity differences and values for \llrnew\ and \llrold\ for
all observed candidate comoving pairs are provided with the associated table
(\tablename~\ref{tbl:data-pairs}; see Appendix~\ref{sec:table}).

\begin{figure}[htb]
  \begin{center}
    \includegraphics[width=0.9\linewidth]{dx-dv.pdf}
  \end{center}
  \caption{%
    Point estimate of the 3D velocity difference between two stars in a comoving
    pair, $|\bs{v}_1 - \bs{v}_2|$, plotted against a point estimate of the 3D
    spatial separation of the two stars, $|\bs{x}_1 - \bs{x}_2|$.
    The distances used in the point estimates are computed using a
    Lutz-Kelker-corrected parallax value (\citealt{Lutz:1973}).
    \label{fig:dx-dv}}
\end{figure}

@@

\subsection{Confirmed comoving pairs}\label{sec:genuine}

\figurename~\ref{fig:dx-dv} shows the magnitude of the 3D velocity difference
plotted against the 3D positional separation of the two stars in all observed
pairs from this work.
We again plot the median of the velocity difference distribution, and here use
the Lutz-Kelker-corrected distance estimate (\citealt{Lutz:1973}) computed from
the stars' \tgas\ parallax measurements to compute the 3D positions of the two
stars.
Small, gray markers are pairs for which $\llrnew < \llrold$, i.e. where the
radial velocity measurements weaken or change our belief that a given pair is
comoving.
Larger, black markers are confirmed comoving pairs ($\llrnew > \llrold$); note
that many large-separation pairs have been ruled-out as comoving, but
most small-separation pairs maintain their status as comoving.

\figurename~\ref{fig:separation} shows histograms of the spatial separation and
sky separation for all pairs; background, gray histogram shows all observed
pairs, and foreground, black histogram shows all pairs for which $\llrnew >
\llrold$.
In this and subsequent figures, we use the phrase ``tangential separation'' (and
symbol $s_{\rm tan}$) to refer to the chord length separation between the two
pairs assuming the two stars are at the same distance (the smaller distance of
the two, $d_{\rm min}$):
\begin{equation}
    s_{\rm sky} = 2\, d_{\rm min} \, \sin{\left( \frac{1}{2}\theta \right)}
\end{equation}
where $\theta$ is the angular extent of the pair.

At sky separations $\gtrsim 1~\pc$, $\approx 40\%$ of the comoving pairs are
confirmed to be comoving; interestingly, the separation distribution of the
genuine pairs appears to be uniform or possibly growing in number with
increasing separation.
This strongly suggests the existence of a population of unbound wide binaries or
dissolved open clusters, which maintain approximate spatial coherence for long
times because of the small relative velocities at the time of disruption of the
constituent stars (e.g., \citealt{Jiang:2010}).
The kinematic properties of these unbound binaries and associations contain
information about the tidal field and fluctuations in the mass distribution of
the Milky Way.
With \gaia\ \DR{2}, we will be able to study the statistical properties of these
widely-separated but comoving stars at even larger separations ($>10~\pc$).

\begin{figure}[htb]
  \begin{center}
    \includegraphics[width=\linewidth]{separation-hist.pdf}
  \end{center}
  \caption{%
    \emph{Left}: Distribution of point estimates of the 3D spatial separation of
    the stars in all observed comoving star pairs (gray), and only those
    confirmed as comoving (black).
    Note that the false-positive rate is $\approx 60\%$ at large separations
    ($>1~\pc$), but there are still a significant number of comoving pairs at
    these very wide separations.
    \emph{Right}: Distribution of tangential separations for all comoving star
    pairs (gray) and only those identified as comoving with 3D velocity
    information (black).
    Noise in the parallax measurements makes the 3D separation distribution
    appear roughly uniform.
    Even the distribution in tangential separation appears to be approximately
    linear for large separations ($\gtrsim 0.1~\pc$).
    \label{fig:separation}}
\end{figure}

\begin{figure}[p]
  \begin{center}
    \includegraphics[width=\linewidth]{separation-with-rave.pdf}
  \end{center}
  \caption{%
    \emph{Top}: Distribution of tangential separation for all confirmed comoving
      star pairs, either from the radial velocity follow-up presented in this
      work or from RAVE \DR{5} radial velocities.
      At separations $\lesssim 10^{-1}~\pc$, the distribution (in log-log)
      appears uniform, consistent with past measurements of the wide binary
      separation distribution.
      At separations between $\approx 10^{-1}$--$10^{0}~\pc$, there appears to
      be an underabundance of comoving pairs, but the number of comoving pairs
      then begins to increase again out to our sample separation limit of
      $10~\pc$.
    \emph{Middle}: Tangential separation plotted against mean distance for each
      comoving pair identified from this work (circles, blue) or from RAVE
      overlap (squares, orange).
    \emph{Right}: Distance distribution for all confirmed comoving pairs.
      Note that at distances $\gtrsim 150~\pc$, our detection efficiency appears
      to decline.
    \label{fig:separation-RAVE}}
\end{figure}

An additional 202 candidate comoving pairs (\citealt{Oh:2017}) have
previously-measured radial velocities (for each star) reported in \DR{5} of the
Radial Velocity Experiment (RAVE; \citealt{Kunder:2017}).
Using the same procedure described in \sectionname~\ref{sec:catalog}, we
identify confirmed comoving pairs using the RAVE velocities for all pairs within
$200~\pc$ and find a similar false-positive rate at large separations ($\approx
38\%$ are confirmed).
Of the \nraveoverlap\ RAVE-overlap candidate comoving pairs (within $200~\pc$)
that meet the above criteria, we find \nravecomoving\ confirmed comoving pairs.
We plot the confirmed comoving pairs identified through radial velocity
follow-up in this work along with the RAVE-confirmed pairs in
\figurename~\ref{fig:separation-RAVE}.
Top histogram shows the tangential separation distribution for the joint
sample, scatter plot shows log-separation and the pair distances, and right
histogram shows the distance distribution of the confirmed pairs.
At larger distances ($\gtrsim 150~\pc$) it appears that the detection efficiency
of our method drops, as expected given the hard cut in log-FML-ratio applied to
define the candidate sample.
Reassuringly, there do not appear to be any trends of the separation
distribution with distance.

\figurename~\ref{fig:genuine-highlighted-cmd} (left) shows the color-magnitude
diagram for all confirmed comoving pairs.
As would be expected if all of the pairs are coeval, the majority of pairs lie
along isochrones in color-magnitude space, with the exception of a few visible
outliers.
For example, there appear to be at least a few main-sequence--giant-branch
pairs, and a few pairs that appear to cross the main-sequence.
We highlight a few of these pairs in the next section
(\sectionname~\ref{sec:interesting-pairs}).

@@

>>>>

\begin{figure}[!ht]
  \begin{center}
    \includegraphics[width=\linewidth]{rv-comparison.pdf}
  \end{center}
  \caption{%
    Comparison of radial velocities measured from our observations with radial
    velocities obtained from the literature search described in
    \sectionname~\ref{sec:rv-validation}.
    Left panel shows our values on the $x$-axis, ${\rm RV}$, against the
    previous literature values on the $y$-axis, ${\rm RV}_{\rm lit}$, for 155
    observed stars (mainly calibration targets) with previously measured radial
    velocities.
    Right panel shows a histogram of the differences, $\Delta{\rm RV} = {\rm RV}
    - {\rm RV}_{\rm lit}$.
    The estimated standard-deviation of the difference distribution implies a
    global systematic uncertainty of $\approx 20~\kms$.
    \label{fig:compare-previous}}
\end{figure}

\section{Methods: Comoving pair probabilities}\label{sec:pair-probs}

@@

Of the 323 observed star pairs, 12 have at least one member with an unreliable
(flagged) radial velocity measurement from the spectroscopic reduction pipeline.
For the remaining \npairsobs\ pairs, we construct a statistical model to
determine the probability that a given candidate comoving pair is comoving in 3D
velocity given the radial velocity and uncertainty measured in this work
combined with \tgas\ astrometric data.
For this inference, the data are the proper motions in the ICRS reference frame,
$\bs{\mu} = (\mu_{\alpha^*}, \mu_\delta)$,\footnote{$\mu_{\alpha^*} = \mu_\alpha
\, \cos\delta$} and parallax, $\varpi$, (from the
\tgas\ catalog) and barycentric radial velocity, $v_r$, for each star in a
candidate comoving pair; we assume that the sky position for each star $(\alpha,
\delta)$ is known perfectly.
For brevity in the expressions below, we define $\bs{\eta} = (\mu_{\alpha^*},
\mu_\delta, v_r)$, where, e.g., $\bs{\eta}_1$ refers to the proper motion and
radial velocity data for star 1 in the pair.
To simplify some expressions, we additionally ignore any correlations between
the observed parallax and the proper motion components (but keep the covariance
between the two proper motion components).

@@

At the smallest tangential separations of our candidate pairs, ($s_{\rm tan}
\approx 10^{-3}~\pc$), if the two stars are bound, the expected velocity
difference is a few $\kms$ (depending on mass).
We use the phrase ``tangential separation'' (and symbol $s_{\rm tan}$) to refer
to the chord length separation between the two pairs assuming the two stars are
at the same distance (the smaller distance of the two, $d_{\rm min}$):
\begin{equation}
    s_{\rm sky} = 2\, d_{\rm min} \, \sin{\left( \frac{1}{2}\theta \right)}
\end{equation}
where $\theta$ is the angular extent of the pair.
We set $\sigma_{\rm co} = 1~\kms$ (rather than using a delta-function at zero)
so that binary stars with very small but finite velocity differences are
considered comoving.
In principle, this ``tolerance'' on finite velocity difference should decrease
with increasing separation, at least out to the radius at which most wide
binaries are disrupted.
At the unbinding radius, the instantaneous velocity difference between the two
stars should be comparable to the orbital velocity of the binary at unbinding
(typically $\approx$0.1--$0.5~\kms$; \citealt{Jiang:2010}), but the difference
may also increase steadily through dynamical heating.
Rather than constructing a model for the full process in this work, we use a
constant velocity tolerance ($\sigma_{\rm co}$) and defer to future work a full
population analysis that also infers the separation distribution of comoving
stars, which could instead express the model as a phase-space distribution
function so that the velocity difference tolerance effectively becomes a
function of separation.
We note that even with this tolerance, it is very improbable that two stars
sampled at random from the field population will have a 3D velocity difference
$<1~\kms$.

\begin{figure}[htb]
  \begin{center}
    \includegraphics[width=0.75\linewidth]{mixture-model-demo.pdf}
  \end{center}
  \caption{%
    Probability distributions (normalized to have peak at 1) used for the
    mixture model described in \sectionname~\ref{sec:pair-probs} as a function
    of true 3D velocity difference, $\psi = \left| \bs{v}_1 - \bs{v}_2 \right|$.
    The field component, $p_{\rm field}(\psi)$, is Maxwellian in form with a
    dispersion parameter $\sigma_{\rm field} \approx 18~\kms$.
    The comoving component, $p_{\rm co}(\psi)$, is a normal distribution
    centered on 0 with a dispersion $\sigma_{\rm co} = 1~\kms$, truncated such
    that $\psi \geq 0$.
    \label{fig:mixture-model-demo}}
\end{figure}

\figurename~\ref{fig:mixture-model-demo} shows the probability distributions for
each mixture model component, $p_{\rm field}(\psi)$ and $p_{\rm co}(\psi)$,
normalized so that the peak of the distribution is at 1.
As noted above, even though the field dispersion is set to $\sigma_{\rm field}
\approx 18~\kms$, most randomly paired stars will have 3D velocity differences
(i.e. $\psi$) significantly larger than the value of the dispersion.

To generate samples in $f$, we use Markov Chain Monte Carlo sampling as
implemented in \package{emcee} (\citealt{Foreman-Mackey:2013}).
We use 28 walkers and keep track of the value of the marginal likelihood under
each mixture model component for each sample at every iteration step.
We run the sampler for 256 burn-in steps from a small ball of initial positions
centered on $f=0.5$, then restart the walkers from the final position and run
for an additional 1024 steps.
With the samples in $f$ and the mixture model component likelihoods, we compute
the per-pair probabilities of belonging to the comoving component (see the final
section of \citealt{Foreman-Mackey:2014}).

@@

\subsection{Probable comoving pairs}\label{sec:genuine}

We use the model described in \sectionname~\ref{sec:pair-probs} to constrain the
fraction of stars consistent with being comoving given 3D velocities for
candidate comoving pairs (a) with radial velocities obtained and described in
this work---\npairsobs\ pairs, and (b) with radial velocities previously
measured and released in \DR{5} of the Radial Velocity Experiment
(\acronym{RAVE}; \citealt{Kunder:2017})---202 pairs.
Using the posterior samples in $f$, we compute the probability (under this
model) that each given pair is comoving including the radial velocity
measurements.
\figurename~\ref{fig:f-samples} shows histograms of posterior pdf samples in $f$
for pairs with our radial velocity measurements (blue, darker histogram) and
with \acronym{RAVE} radial velocities (orange, lighter histogram); in either
case, $f \approx 40\%$ implying a contamination rate $1-f \approx 60\%$.

\begin{figure}[htb]
  \begin{center}
    \includegraphics[width=0.7\linewidth]{f_samples.pdf}
  \end{center}
  \caption{%
    Histograms of samples in $f$, the fraction of comoving pairs, for candidate
    pairs with radial velocities from this work (blue, darker histogram) and
    from \acronym{RAVE} (orange, lighter histogram).
    For the \npairsobs\ candidate pairs observed in this work, this amounts to
    $\approx$120 comoving pairs.
    For the 202 candidate pairs in \acronym{RAVE}, this amounts to $\approx$90
    comoving pairs.
    \label{fig:f-samples}}
\end{figure}

\figurename s~\ref{fig:dx-dv} and \ref{fig:dx-dv-rave} show the magnitude of the
3D velocity difference plotted against the tangential separation of the two
stars in all candidate pairs with radial velocities measured in this work and
from \acronym{RAVE}, respectively.
The error bars show the 15--85th percentiles of the 3D velocity difference
distributions for each candidate pair and black points mark the medians.
The velocity difference distributions for true comoving pairs with radial
velocity precisions of $\approx5~\kms$ often have peaks at smaller velocity
differences, but have medians at larger values because the distributions have
large tails.
This is visible in the plot in that the lower error bars tend to be smaller than
the upper error bar, indicating that the distributions tend to be more
concentrated towards smaller velocities.
This, combined with the fact that it is very improbable to get 3D velocity
differences around $10~\kms$ from the field mixture component, ultimately means
that many of the pairs with median 3D velocity differences close to $\approx
10~\kms$ actually end up having a high probability of being comoving (as is seen
in the blue points in these figures).

\begin{figure}[htb]
  \begin{center}
    \includegraphics[width=0.9\linewidth]{dx-dv-apw.pdf}
  \end{center}
  \caption{%
    Three-dimensional velocity difference between two stars in a comoving pair,
    $|\bs{v}_1 - \bs{v}_2|$, plotted against a point estimate of the tangential
    (sky) separation of the two stars, $s_{\rm tan}$.
    Each point corresponds to one candidate comoving pair and is colored by the
    probability of being comoving under the mixture model described in
    \sectionname~\ref{sec:pair-probs}.
    Error bars in velocity represent the 15--85th percentiles of the inferred
    3D velocity differences for each pair.
    The left panel shows small-separation pairs ($s_{\rm tan} < 1~\pc$) in
    log-space; the right panel shows large-separation pairs ($1 < s_{\rm tan} <
    10~\pc$) with a linear horizontal axis.
    \label{fig:dx-dv}}
\end{figure}

\begin{figure}[htb]
  \begin{center}
    \includegraphics[width=0.9\linewidth]{dx-dv-rave.pdf}
  \end{center}
  \caption{%
    Same as \figurename~\ref{fig:dx-dv} for pairs with radial velocities
    measured by the \acronym{RAVE} survey.
    \label{fig:dx-dv-rave}}
\end{figure}

\subsection{Separation distributions}\label{sec:separations}

\figurename~\ref{fig:separation} shows histograms of the tangential separation
for all pairs with radial velocities from this work and the \acronym{RAVE}
survey, with each pair weighted by the probability that it is comoving.
Interestingly, the separation distribution of the probable comoving pairs
appears to be uniform with increasing separation.
This strongly suggests the existence of a population of unbound wide binaries or
dissolved open clusters, which maintain approximate spatial coherence for long
times because of the small relative velocities at the time of disruption of the
constituent stars (e.g., \citealt{Jiang:2010}).
The kinematic properties of these unbound binaries and associations contain
information about the tidal field and fluctuations in the mass distribution of
the Milky Way.
With \gaia\ \DR{2}, we will be able to study the statistical properties of these
widely-separated but comoving stars at even larger separations ($>10~\pc$).

\begin{figure}[htb]
  \begin{center}
    \includegraphics[width=\linewidth]{separation-hist.pdf}
  \end{center}
  \caption{%
    Weighted histogram---in log-log (\emph{left panel}) and linear (\emph{right
    panel}) scales---of the tangential separation of all comoving star pairs
    with radial velocity measurements either from this work or from the
    \acronym{RAVE} survey.
    At separations $\lesssim 10^{-1}~\pc$, the distribution (in log-log) appears
    uniform, implying a decreasing distribution in linear space that is
    qualitatively consistent with past measurements of the wide binary
    separation distribution.
    At separations between $\approx 10^{-1}$--$10^{0}~\pc$, there appears to be
    an underabundance of comoving pairs, but $\gtrsim 1~\pc$ the number of
    comoving pairs appears to be uniform out to our sample separation limit of
    $10~\pc$.
    \label{fig:separation}}
\end{figure}

To see whether the separation distribution depends on heliocentric distance, we
also plot the tangential separation of all comoving pairs with comoving
probability $> 50\%$ as a function of heliocentric distance in
\figurename~\ref{fig:separation-scatter}.
The top histogram again shows the tangential separation distribution for the
joint sample (now only for pairs with $> 50\%$ comoving probability).
The scatter plot shows tangential separation and the pair distances, and
the right histogram shows the distance distribution of these high-probability
pairs.
Reassuringly, there do not appear to be any trends of the separation
distribution with distance.

\begin{figure}[htb]
  \begin{center}
    \includegraphics[width=0.9\linewidth]{separation-with-rave.pdf}
  \end{center}
  \caption{%
    \emph{Top}: Distribution of tangential separation for all comoving star
      pairs with probability $>50\%$, either from the radial velocity follow-up
      presented in this work or from \acronym{RAVE} \DR{5} radial velocities.
    \emph{Middle}: Tangential separation plotted against mean distance for each
      comoving pair identified from this work (circles, blue) or from RAVE
      overlap (squares, orange).
    \emph{Right}: Distance distribution for all confirmed comoving pairs.
      Note that at distances $\gtrsim 150~\pc$, our detection efficiency appears
      to decline.
    \label{fig:separation-scatter}}
\end{figure}

\figurename~\ref{fig:genuine-highlighted-cmd} (left) shows the color-magnitude
diagram for pairs with radial velocities measured in this work with comoving
probability $> 0.5$.
As would be expected if all of the pairs are coeval, the majority of pairs lie
along isochrones in color-magnitude space, with the exception of a few visible
outliers.
For example, there appear to be at least a few main-sequence--giant-branch
pairs, and a few pairs that appear to cross the main-sequence.
We highlight a few of these pairs in the next section
(\sectionname~\ref{sec:interesting-pairs}).

@@

% \section{Old section on FMLs}
% We compute two fully marginalized likelihoods (FMLs) for the data: (1) assuming
% that the two stars are comoving with identical 3D velocity vectors (within some
% allowed small dispersion), and (2) assuming that the two stars have independent
% 3D velocity vectors drawn from a prior velocity distribution.
% We set the small allowed dispersion ($s$, in previous work) to $1~\kms$, and use
% a mixture of three Gaussian velocity components for the prior, meant to
% represent the thin disk, thick disk, and halo velocity distributions relative to
% the sun (see \citealt{Oh:2017}).

% To compute the new likelihood ratio, we modify the expressions from
% \citet{Oh:2017} to include the radial velocity measurements.
% In the same notation, this requires the following changes:
% \begin{itemize}
%   \item The projection matrix, $\mat{M}$, now has to include the radial unit
%     vector
%     \begin{equation}
%       \mat{M} =
%         \left(
%           \begin{array}{c c c}
%             -\sin\alpha & \cos\alpha & 0 \\
%             -\sin\delta \, \cos\alpha & -\sin\delta \, \sin\alpha & \cos\delta\\
%             \cos\delta \, \cos\alpha & \cos\delta \, \sin\alpha & \sin\delta
%           \end{array}
%         \right)
%     \end{equation}
%     where $(\alpha, \delta)$ are the ICRS coordinates (right ascension and
%     declination) of a given star.
%   \item The data-space velocity vector, $\vec{y}$, for one star now becomes
%     \begin{equation}
%       \vec{y} =
%         \transpose{\left(
%           \begin{array}{c c c}
%             r_i\,\mu_{\alpha}^* &
%             r_i\,\mu_{\delta} &
%             v_{r}
%           \end{array}
%         \right)}
%     \end{equation}
%     where $\mu_\alpha^*$ is the proper motion in right ascension (including the
%     $\cos{\delta}$ term), $\mu_\delta$ is the proper motion in declination, and
%     $v_r$ is the radial velocity.
%   \item The covariance matrix becomes the block-diagonal matrix
%     \begin{equation}
%       \mat{\Sigma} = \left(
%         \begin{array}{c c}
%           r^2 \, \mat{C} & 0 \\
%           0 & \sigma_{v_r}^2
%         \end{array}
%       \right)
%     \end{equation}
%     where $r$ is the distance, $\mat{C}$ is the proper motion covariance matrix
%     (from the \tgas\ catalog), and $\sigma_{v_r}$ is the radial velocity
%     uncertainty.
% \end{itemize}
% To compute the FML for hypothesis 1 for a pair of stars $(i,j)$, the data-space
% vector must become the 6-vector
% \begin{equation}
%   \vec{y} = \transpose{\left(
%     \begin{array}{c c}
%       \vec{y}_i &
%       \vec{y}_j
%     \end{array}
%     \right)}
% \end{equation}
% and the covariance matrix becomes the block-diagonal matrix
% \begin{equation}
%   \mat{\Sigma} = \left(
%     \begin{array}{c c}
%       \mat{\Sigma}_i & 0 \\
%       0 & \mat{\Sigma}_j
%     \end{array}
%   \right) \quad .
% \end{equation}

% \begin{figure}[htbp]
%   \begin{center}
%     \includegraphics[width=0.5\linewidth]{llr-vs-dv.pdf}
%   \end{center}
%   \caption{%
%     Point estimate of the 3D velocity difference between two stars in a comoving
%     pair, $|\bs{v}_1 - \bs{v}_2|$, plotted against the change in
%     log-likelihood-ratio computed when including radial velocity data.
%     Error bars are shown for a few randomly-chosen, representative pairs and
%     represent the 15th and 85th percentiles of velocity differences computed
%     from samples from the error distributions of parallax, proper motion
%     components, and radial velocity.
%     The likelihood ratio increases (i.e. the belief that the stars are comoving
%     becomes stronger) when the 3D velocity difference is small.
%     \label{fig:llr}}
% \end{figure}

% If the new likelihood ratio is larger than that computed from astrometry alone,
% the added radial velocity information supports the belief that the two stars are
% truly comoving in 3-space.
% Conversely, if the new likelihood ratio is smaller, the radial velocity data
% indicate that the pair is likely a false-positive and are not truly comoving.
% In practice, we compute the logarithm of the likelihood ratios,
% $\mathcal{R} = \ln \mathcal{L}_1/\mathcal{L}_2$; we refer to the
% proper-motion-only log-FML-ratio for a given pair of stars as
% \llrold\ (as computed in \citealt{Oh:2017}) and the log-FML-ratio including the
% radial velocity data as \llrnew.
% We consider a pair as being genuinely comoving if $\llrnew > \llrold$:
% \ncomoving\ of the \npairsobs\ comoving pairs satisfy this constraint.
% \figurename~\ref{fig:llr} shows the magnitude of the 3D velocity difference
% plotted against the change in log-likelihood-ratio, $\llrnew - \llrold$, for all
% observed pairs.
% In detail, for visualization, we sample from the error distributions over proper
% motion, parallax, and radial velocity for each star individually and compute the
% magnitude of the 3D velocity differences, $|\bs{v}_1-\bs{v}_2|$, for $2^{16}$
% samples.
% We plot the median of the resulting distribution of velocity-differences as
% dark markers (circles).
% The error bars show the 15th and 85th percentiles of the distributions for a few
% randomly chosen pairs.
% Note that pairs with large 3D velocity differences have large (negative) changes
% to their log-likelihood-ratio values.
% Some pairs with large velocity differences also have small (near zero) changes
% to the log-likelihood ratio: these are pairs where each member has a velocity
% that is an outlier with respect to the prior velocity distribution.

% The inferred 3D velocity differences and values for \llrnew\ and \llrold\ for
% all observed candidate comoving pairs are provided with the associated table
% (\tablename~\ref{tbl:data-pairs}; see Appendix~\ref{sec:table}).

% \begin{figure}[htb]
%   \begin{center}
%     \includegraphics[width=0.9\linewidth]{dx-dv.pdf}
%   \end{center}
%   \caption{%
%     Point estimate of the 3D velocity difference between two stars in a comoving
%     pair, $|\bs{v}_1 - \bs{v}_2|$, plotted against a point estimate of the 3D
%     spatial separation of the two stars, $|\bs{x}_1 - \bs{x}_2|$.
%     The distances used in the point estimates are computed using a
%     Lutz-Kelker-corrected parallax value (\citealt{Lutz:1973}).
%     \label{fig:dx-dv}}
% \end{figure}
